Introducing Statistical Design of Experiments to SPARQL Endpoint Evaluation

### 2013-05-05 exp1/iswc2013.tex
Small changes Note warm-up and caching
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@507 43ef1190-90fd-4dd7-a769-d0121f02b80e

Thus, attacking the complexity problem by eliminating parameters we
think influence the outcome for simplicity is indefensible as
scientific practice. We must instead tackle this problem by enhancing
our capability of testing complex systems and to identify situations
where our assumptions are flawed. Such constraints can only be made
when we have very good reason to suspect that a certain parameter does
not have influence.

@@

Then, we demonstrate how a fractional factorial experiment can be used
to maintain a structured approach to the problem while reducing the
number of runs the experiment needs, and we discuss the tradeoffs
involved. We then continue to discuss techniques that shows our
simplistic experiment to be flawed, yet scientifically better than the
practice of benchmarking. Finally, we outline a roadmap to establish
the use of these techniques in SPARQL endpoint evaluations.

@@

To only deal with the relatively straightforward aspects of the DoE
formalism, we have chosen to constrain the experiment strictly to 2
levels. We have chosen 8 factors each having 2 levels, ``TripleC'' the number of triples in the
dataset, 1 or 2 MTriples, ``Machine'', which is the software and hardware platform,
where one machine runs GNU/Linux Debian Squeeze, has 16 GB RAM, an
Intel Core2 Duo E8500 CPU and two Parallell-ATA disks in a RAID-1
configuration. The other machine runs Debian Wheezy, has 8 GB RAM, an
Intel Core i7-3520M CPU and a single SSD on SATA-III.

@@

These fragments has been carefully designed for illustrative utility,
as well as suitable selectivity. When the experiment is run, they are
combined to yield a query, which is then sent to a SPARQL
endpoint. The SPARQL endpoint is itself set up using 4store\todo{ref}
version 1.1.5. The ``Machine'' factor specifies a hostname, whereas
``TripleC'' and ``Implement'' specifies a port number. On each host,
four 4store instances runs independently on different
ports. \todo{possible problem?}

@@

>>>>

Thus, attacking the complexity problem by eliminating parameters we
think influence the outcome for simplicity is indefensible as
scientific practice. We must instead tackle this problem by enhancing
our capability of testing complex systems and to identify situations
where our assumptions are flawed. Such constraints can only be made
when we have very good reason to suspect that a certain parameter does
not have any influence.

@@

Then, we demonstrate how a fractional factorial experiment can be used
to maintain a structured approach to the problem while reducing the
number of runs the experiment needs, and we discuss the tradeoffs
involved. We then continue to discuss techniques that shows our
simplistic experiment to be flawed, yet scientifically better than the
practice of benchmarking. Finally, we outline a roadmap to establish
the use of these techniques in SPARQL endpoint evaluations both as a
practical tool and as a scientific more rigorous methodology.

@@

First, note that the concept of parameters as considered in the
introduction is known as \emph{factors} in DoE. For each factor, we
have chosen in this paper to only deal with the relatively
straightforward aspects of the DoE formalism, and therefore 
constrain the experiment strictly to 2 levels.

We have chosen 8 factors each having 2 levels, ``TripleC'' the number of triples in the
dataset, 1 or 2 MTriples, ``Machine'', which is the software and hardware platform,
where one machine runs GNU/Linux Debian Squeeze, has 16 GB RAM, an
Intel Core2 Duo E8500 CPU and two Parallell-ATA disks in a RAID-1
configuration. The other machine runs Debian Wheezy, has 8 GB RAM, an
Intel Core i7-3520M CPU and a single SSD on SATA-III.

@@

These fragments has been carefully designed for illustrative utility,
as well as suitable selectivity. They are not themselves important,
they only serve to illustrate certain examples of possible
factors. When the experiment is run, they are combined to yield a
query, which is then sent to a SPARQL endpoint. The SPARQL endpoint is
itself set up using 4store\todo{ref} version 1.1.5. The ``Machine''
factor specifies a hostname, whereas ``TripleC'' and ``Implement''
specifies a port number. On each host, four 4store instances runs
independently on different ports. \todo{possible problem?}

@@

Finally note that there are two common issues in benchmarking we do
not consider: we do not allow a warm-up run, nor do we take into
account that the server may cache all or parts of the result set. The
reasons for this choice will be discussed later.\todo{lurking
  variables}
